### 2021-05-25 paper/paper.tex

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath, enumitem, url}
\usepackage{xcolor}

% For figures.
% Package float, adds the H option which forces figure placement.
\usepackage{graphicx, float}
\graphicspath{ {./figures/} }

\newcommand\todo{\textcolor{red}{\textbf{[TODO:] }}}

\title{CSE571 Project 2}
\author{Group: Dropouts}
\date{June 2021}

\begin{document}
\maketitle

\section{Group Information}

Members:
\begin{itemize}
    \item Keyan Pishdadian (keyanp@cs.washington.edu)
    \item Jakub Filipek (balbok@cs.washington.edu)
    \item Kyle Deeds (kdeeds@cs.washington.edu)
\end{itemize}

\noindent Source code for behavior cloning approach:
\newline
\url{https://github.com/keyan/duckiebot_behavior_cloning}

\noindent Source code for reinforcement learning approach:
\newline
\url{https://github.com/balbok0/cse571-sp21-project-2-dropouts}

\section{Task Overview}

\todo Reproducing RL research usually doesn't work: https://arxiv.org/abs/1709.06560

\section{Environment Setup}
\subsection{Simulation}
\todo Discuss gym-duckietown and AIDO submissions

\subsection{Real World}
\todo Show different maps used

% \begin{figure}[H]
% \includegraphics[width=8cm,keepaspectratio]{map_simple_loop}
% \centering
% \caption{Photo of physical layout of map 1.}
% \label{fig:map_simple_loop}
% \end{figure}

\section{Reinforcement Learning Approach}
\todo Discuss our initial approach using RL

\todo how we anticipated this would work

\todo ultimately how it performed poorly

\subsection{DQN}
\todo Discuss general algorithmic idea behind DQN

\todo Show training results using DQN and videos of simulator performance

\subsection{DDPG}
\todo Discuss general algorithmic idea behind DDPG

\todo Show training results using DDPG and videos of simulator performance

\section{Behavior Cloning Approach}
\todo High level overview of behavior cloning and end-to-end control with monocular images

\subsection{Data Preparation}
\todo Discuss data collection procedure and cleaning

\todo Discuss final datasets used

\subsection{Model Architectures}
\todo Discuss each model architecture evaluted

\todo Show data from training

\section{Performance Evaluation}
\subsection{Simulator and AIDO}
\todo Share links to videos from simulator and AIDO submissions

\todo Discuss quantative performance for each model including baseline

\subsection{Real World}

\todo Discuss quantative performance for each model including baseline

\todo Share links to videos for each map/model evaluated

\section{Potential Future Improvements}

\todo Discuss algorithmic enhancements

\todo Discuss robot limitations

\bibliographystyle{alpha}
\bibliography{paper}

\end{document}